---
title: "Aides Drug"
author: "Adrian Cruz"
date: "October 8 2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new  preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the females, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

# Methods


$H_0$: The vaccine is not effective in preventing women from contracting aids.  
$H_A$: The vaccine is effective in preventing women from contracting aids.  
infected~gender   Cat~Cat
<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


# Results
<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->
```{r}
women <- triUmp[triUmp$gender=='F',]
```


## Descriptive Results

### Graphical Descriptive Results


```{r}
barchartGC(~infected + treatment,data=women)
barchartGC(~infected + treatment,data=women, type="percent")

```

For women who contracted aids, most of them had undergone treatment. Therefore the treatment does not appear to be effective in preventing aids but rather contracting aids.
<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

### Numerical Descriptive Results

```{r}
table1 <- xtabs(~infected + treatment, data=women)
rowPerc(table1)
colPerc(table1)
```

Among those who got aids, 58.82% of them recieved the treatment.

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

With this data, we must reject the null hypthesis because the P-value is lower than 0.05 with the p value being 7.686e-05.
The odds show that if you are a women and you take the treatment, you are 1.54 times more likely to attract aids than those who didn't.

# Discussion and Conclusion

In this assignment, we conducted several tests to prove that either the treatment is effective or not effective in preventing aids in women. During a graphical test, we were shown that women who orginally took the treament actually contracted aids rather than preventing them. This was also shown to be true through numerical results with 58.82% of those who had aids, actually recieving the treatment. Finally, we can reject the null hypothesis by viewing that the P-value was 6.596e-05, which was below 0.05. Therefore the vaccine is not effective in preventing women from contracting aids.

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
